Research Experience

### 2010-06-23 doc/thesis/thesis.tex
added some bullets
more work on the thesis.tex, kalman filter stuff

\usepackage{times}
%\usepackage{tikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{cancel}
\usepackage{hyperref}

@@

@@

The goal of the particle filter is to evolve a probably distribution 
$Pr(\hat{x}(T) | u[0:T], y[0:T])$,
that asymptotically approaches the probability distribution $Pr(x(T) | u[0:T])$.
Considering that $y$ contains measurement noise and noise in $x$ can drive
changes in $y$, it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
but a true posterior. 
To begin with, the particle filter starts with a proposal distribution, and $N_p$
particles need to be drawn from that distribution, $\alpha(x)$:
\begin{equation}
\{\hat{Pr}x_i(0),w_i] : x_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
\end{equation}

@@

@@

For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
\{100, 500, 1000\} particles
\begin{enumerate}
\item Comparison with a linear system with similar number of degrees of freedom
\item Ground truth vs. Estimated signal during particle filter run
\item Ground truth vs. Estimated signal with final parameter set
\item True Parameters vs. Final Parameter Sets
\item Variance of final parameters when faced with same ground truth, different noise
\item Variance of final parameters when faced with same ground truth, different noise
\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
\item Estimator Variance based on different noise runs
\item Final Particle Distribution
\end{enumerate}

>>>>

\usepackage{times}
\usepackage[siunitx]{circuitikz}
%\usetikzlibrary{trees}
%\usetikzlibrary{shapes,snakes}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithmic}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{cancel}
\usepackage{hyperref}

@@

\section{Unscented Kalman Filter}
The unscented kalman filter is a powerful bayes-gaussian filter that attempts
to model the posterior distribution of dynamical systems as a multivariate
Gaussian. The Unscented Kalman Filter (UKF) generalizes the Extended Kalman
Filter by allowing the state update to be a function, $g$,

\begin{eqnarray}
X(t) = g(u(t), X(t-1))\\
Y(t) = h(X(t)
\end{eqnarray}

In order to estimate the posterior at $t$, a deterministic set of points (often 2 
per dimension, plus 1 at the mean)
weighted according to a Gaussian estimate of $X(t-1)$ are passed through
the update equation. This set of points are then used to estimate the 
mean and covariance of $X(t)$. The benefit of this, is that it requires
no Jacobian and only a few extra calculations to get a decent esimate of
a posterior Gaussian. In the BOLD case, the set of equations we are modeling have no 
closed form solution, and finding the Jacobian is impossible without approximations. Although 
\cite{Riera2004} claims to use a Jacobian for their solution, they are
in fact using the Jacobian of $\dot{s}$, $\dot{q}$, $\dot{f}$, $\dot{v}$,
rather than $s$, $q$, $f$, and $v$. This is important because the Extended
Kalman filter depends on the jacobian to map a gaussian through the 
advancement of time. Thus the popular Extended Kalman Filter won't work
in this case, whereas the Unscented Kalman Filter still does.

The difficulty with the Kalman filter is that it is not well suited
to parameter estimation. If we wanted to find the mean of a static
parameters, such as $V_0$, it would be necessary to guess a mean 
and variance for all time. This is because the Kalman filter is primarily
concenrned with estimating the output, and is not really concerned
with estimating the underlying state, $X(t)$. As such estimated 
output, $\hat{Y}(t)$, is simply a weighted sum of the measurement,
$Y(t)$, and the expected value of the output based on the propagated
state variables, $\hat{X}(t)$. As such, when the  the filter will attenuate measurements
the vastly disagree with $E[h(\hat{X}(t))]$, but eventually

The difficulty
of using a Kalman Filter, however, is that it assumes a multivariate 
Gaussian for the state variables, $X(t-1)$. The problem with this is that 
when a system is significantly nonlinear, the state at 
$X(t)$ will almost certainly be non-Gaussian, and thus estimating
$X(t+1)$ with $X(t)$ as a multivariate Gaussian will perpetually introduce
error in the distribution. Furthermore, it is not really known what 
sort of underlying distributions may exist in such a mixed biological,
mechanical, chemical system such as the brain. Assuming the parameters
all to be Gaussian may in fact be a gross error. On the other hand, if
the our tests indicate a Gaussian distribution is a good fit, it would
be a wise plan to go back and consider the use of a modified Unscented 
Kalman Filter rather
than the full Particle Filter we discuss in the next section.

@@

The goal of the particle filter is to evolve a probably distribution 
$Pr(\hat{x}(T) | u[0:T], y[0:T])$,
that asymptotically approaches the probability distribution $Pr(x(T) | u[0:T])$.
Considering that $y$ contains measurement noise as well as noise from $x$,
 it is clear that $Pr(x(t) | u[0:T])$ is not a single true value
but a true posterior. 
To begin with, the particle filter starts with a prior distribution, and $N_p$
particles need to be drawn from that distribution, $\alpha(x)$:
\begin{equation}
\{\hat{Pr}x_i(0),w_i] : x_i(0) \sim \alpha(x), w_i = \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}
\end{equation}

@@

\section{Simple Example}
So a simple example may be used to demonstrate the proposed use of a particle filter.
Thus, let us consider a simple half wave rectifier circuit, shown in figure 
\ref{fig:HalfWaveRectifier}

\begin{figure}
\begin{center}
\begin{circuitikz}[scale=2, american]
\draw
 (0,0)  node[transformer core] (T) {}
 (T.A1) -- (-1,0)
 (T.A2) -- (-1,-1.05)  to[V, v=$u(t)$] (-1, 0)
 (T.B1) -- (.5, 0) to[D, l=$v_t$] (1.5,0) to[C=$C$] (1.5, -1.05)
 (1.5, 0) -- (2.5, 0) to[R=$Rm$, v=$v_y$] (2.5, -1.05) -- (T.B2) 
 (T.base) node {G};
\end{circuitikz}
\label{fig:HalfWaveRectifier}
\caption{An Example Half Wave Rectifier Circuit, where $G$ is the transformer
gain, $v_t$ is the activation voltage of the diode, $u(t)$ is the input at time $t$, 
$C$ is the capacitance, $R$ is the load resistance and $v_y$ is the output voltage}
\end{center}
\end{figure}

Thus when $u(t)G$ is less than $v_t$, the circuit will discharge the capacitor,
but when $u(t)G$ is greater than $v_t$, the capacitor will charge up. Unlike
the normal rectifier circuit, we'll deal with a generic input
Let $V_C$ be the

@@

For simulated data, single timeseries:

For \{delta, DC/Spline\}, \{exponential, gaussian, cauchy\}, \{biased, unbiased initial\},
\{100, 500, 1000\} particles
\begin{enumerate}
\item Ground truth vs. Estimated signal during particle filter run
\item Ground truth vs. Estimated signal with final parameter set
\item True Parameters vs. Final Parameter Sets
\item Variance of final parameters when faced with same ground truth, different noise
\item MSE of (a new timeseries based on X(t) vs. ground truth) for all t
\item Estimator Variance based on different noise runs
\item Final Particle Distribution
\end{enumerate}

For Simulated Data, Full Volume:
